Exploring dimensions of biodiversity in Japanese ferns

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides

- Combine with broad sampling from genbank (ca. 60 genes x 4,000 taxa) (Nitta et al., in prep.)

@@

- Date with 26 fossil calibration points `r AutoCite(bib, "Testo2016a")` using treePL `r AutoCite(bib, "Smith2012")`

@@

```{r richness, out.width = "120%"}
# Fern richness
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "No. taxa") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  ) +
  ylim(ja_lat_min, ja_lat_max)

@@

```{r ses-pd, out.width = "120%"}
# SES PD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r ses-rpd, out.width = "120%"}
# SES RPD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r ses-rpd-rfd, out.width = "120%"}
# SES RPD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RFD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r pe, out.width = "120%"}
# PE
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla", labels = scales::percent) + 
  labs(fill = "PE percentile") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  ylim(ja_lat_min, ja_lat_max) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

## Conclusions

How is biodiversity distributed?

@@

- Phylogenetic and functional diversity, phylogenetic endemism highest in Okinawa

How is biodiversity structured?

- Into four major clusters that track major forest types

How well is biodiversity protected?

@@

- May be in danger from herbivory by deer

## Future directions

- Understand the role that hybridization has on diversification (ca. 400 described hybrids!)

- Clarify the role of fern gametophytes in community assembly at broad scales

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 1, end = 5)
```
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 6, end = 10)
```
]

>>>>

- Combine with broad sampling from GenBank

@@

- Date with 27 fossil calibration points `r AutoCite(bib, "Testo2016a")` using treePL `r AutoCite(bib, "Smith2012")`

@@

```{r richness, out.width = "120%"}
# Fern richness
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") + 
  labs(fill = "No. taxa") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r ses-pd, out.width = "120%"}
# SES PD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r ses-rpd, out.width = "120%"}
# SES RPD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r ses-rpd-rfd, out.width = "120%"}
# SES RPD
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RFD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

```{r pe, out.width = "120%"}
# PE
# - map
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla", labels = scales::percent) + 
  labs(fill = "PE percentile") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme_pp() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

@@

# Conclusions

- How is biodiversity distributed?

- How is biodiversity structured?

- How well is biodiversity protected?

# Conclusions

```{r conclusions-plots}
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(subtitle = "Richness") +
  simple_map_theme()

b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  labs(subtitle = "RPD") +
  simple_map_theme()

c <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  labs(subtitle = "Taxonomic bioregions") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  simple_map_theme()
```

.pull-left[
How is biodiversity distributed?

@@

- Phylogenetic and functional diversity, phylogenetic endemism highest in Okinawa
]

.pull-right[
```{r conclusions-richness, out.width="450px"}
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(title = "Richness") +
  simple_map_theme() +
  theme(
    plot.title = element_text(size = 18)
  )

b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  labs(title = "RPD") +
  simple_map_theme() +
  theme(
    plot.title = element_text(size = 18)
  )

a + b + plot_layout(ncol = 1)
```
]

# Conclusions

.pull-left[
How is biodiversity structured?

- Into four major clusters that track major forest types
]

.pull-right[
```{r conclusions-regions}
ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  labs(title = "Taxonomic bioregions") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  simple_map_theme_light() +
  theme(
    plot.title = element_text(size = 18)
  )
```
]

# Conclusions
.pull-left[
How well is biodiversity protected?

@@

- May be in danger due to herbivory by deer
]

.pull-right[
```{r conclusions-protection}
 ggplot() +
  geom_sf(data = japan_cropped_to_richness) +
  geom_sf(data = protected_richness, aes(fill = status)) + 
  scale_fill_viridis_d(breaks = c("low", "medium", "high"), labels = c("Low", "Medium", "High")) +
  labs(title = "Top 5% richness hotspots", fill = "Conservation\nstatus"
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 18)
  )
```
]

---
# Future directions

.pull-left[
<br>
- Understand the role that hybridization has on diversification <br> (ca. 300 described hybrids!) `r NoCite(bib, "Ebihara2009")`

<br>
<br><br>
<br>
- Clarify the role of fern gametophytes in community assembly at broad scales
]

.pull-right[
<img src="images/vandenboschia.png" alt="drawing" width="280"/>

<img src="images/IMGP5729.jpeg" alt="drawing" width="300"/><br>.tiny[Photo: Joel Nitta]
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 1, end = 6)
```
]

@@

.small[
```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
PrintBibliography(bib, start = 7, end = 12)
```
]
